---
title: "How to use the functions in this package"
author: "Amrit Singh"
date: "December 3, 2016"
output: 
  html_document: 
    keep_md: yes
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, message = FALSE)
```

```{r loadLibraries}
library(amritr)
library(mixOmics)
```

#biomarker pipeline()
## y~X, y = binary response, X = nxp dataset
### Description of classification algorithms
  * Enet panels (alpha = 0-0.9, step = 0.1) with and without a p-value filter (topranked = # of top ranked significant features to use)
  * random forest panels with and without a p-value filter (topranked = # of top ranked features to use)
  * support vector machine panels with and without a p-value filter (topranked = # of top ranked features to use)
  * single (p) biomarkers (based on a glm model)
  * pathway biomarkers based on geneset from BioCarta_2016, KEGG_2016, Reactome_2016, and WikiPathways_2016 from Enrichr (each panel is a geneset from a given database)

### Inputs and outputs
  * Input:
          + X = nxp dataset
          + y = n-vector of class labels
          + topranked = 30 (number of significant features to input into classifier)
          + validation = "Mfold" or "loo"
          + M = 2 - # of folds in the cross-validation scheme
          + iter - # of times to repeat cross-validation
          + threads - # of cores, each running a cross-validation scheme
          + pathways - list of lists each for a different database; each element is a genset consisting of gene symbols
  * Output:
          + a dataframe with columns, Mean and SD of CV-AUC, Panel_label, Genes (within panels), Type of Panel (Enet, RF, SVM, GLM, Pathways(biocarta, kegg, reactome, wikipathways))

```{r biomarkerPipeline}
data("pathways"); data("breast.TCGA")

Y = breast.TCGA$data.train$subtype
names(Y) <- rownames(breast.TCGA$data.train$mrna)

Y = c(Y[Y == "Basal"][1:10], Y[Y == "Her2"][1:10])
Y[Y == 1] <- "Basal"
Y[Y == 2] <- "Her2"
Y <- factor(Y)
set.seed(123)
X = breast.TCGA$data.train$mrna[names(Y), sample(1:200, 30)]

## run biomarker pipeline for a binary response
allPanels <- biomarkerPipeline(X = X, Y = Y, topranked = 30, validation = "Mfold", M = 2, 
             iter = 2, threads = 2, progressBar = TRUE, pathways = pathways)
head(allPanels)
```

### Plot AUC (Mean +/- SD) of all panels

```{r, fig.height=5, fig.width=8}
allPanels %>% arrange(Mean) %>% mutate(Panel = 1:nrow(.)) %>% 
  ggplot(aes(x = Panel, y = Mean, color = Type)) + geom_point() +
  geom_errorbar(aes(ymin = Mean-SD, ymax = Mean+SD)) +
  customTheme(sizeStripFont = 10, xAngle = 0, hjust = 0.5, vjust = 0.5,
    xSize = 10, ySize = 10, xAxisSize = 10, yAxisSize = 10) +
  ylab("AUC - 2x2-fold CV") + xlab("Panels") +
  geom_hline(yintercept = 0.5, linetype = "dashed")
```

# Concatenation-Enet biomarker panel
### Description of algorithm
  * all X datasets are combined column-wise and a single-dataset classifer is applied (Enet)

```{r}
combinedDat <- do.call(cbind, breast.TCGA$data.train[-length(breast.TCGA$data.train)])
Y.train <- breast.TCGA$data.train$subtype
# Enet
result <- enet(X = combinedDat, Y = Y.train, alpha=1, family="multinomial", lambda = NULL, X.test = NULL, Y.test = NULL,
  filter = "none", topranked = 50)
lapply(breast.TCGA$data.train[-length(breast.TCGA$data.train)], function(i){
  intersect(colnames(i), result$enet.panel)
})

```

